---
title: "Initial Korean covid 19"
author: "Bella Verdin"
date: "Spring 2024"
output:
  html_document: default
  pdf_document: default
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
options(digits=2)
require(tidyverse)
require(tigerstats)
require(lubridate)
kRdata <- read.csv(file = 'kRdata.csv')
kRdata <- kRdata %>%
mutate(across(where(is.character), as.factor))

```
### Data is from the Patient_Info dataset from South Korea 1 Jan to 1 May 2020
### This data was retrieved from a downloadable zip file from kaggle

Some of the data was uncoded for sex - sex must be just male or female
77 of the data entries were coded sex  as blank, so we omit these.
state of a patient must be died or alive - we adjust that as well   
  
# Your Question - Is dying independent of sex?

#Introduction

$H_0$: There is no difference in death rate based on gender. 

$H_A$: There is a difference in death rate based on gender.

#Methods
The variabls for this study are Sex and results 

I want to find out if the results depend on sex. Results are a categorical variable as well as sex.

So I can use the Cat~Cat helper file 

#Results 

## Graphical results - use ggplot commands to do this  


```{r}
barchartGC(~result + Sex,data=kRdata)
barchartGC(~result + Sex,data=kRdata, type="percent")
```

  

With this chart I can see those who died, 60% of the chart were males and females the other 40%



  
## Numerical results

```{r}
table1 <- xtabs(~result + Sex, data=kRdata)
rowPerc(table1)
colPerc(table1)
```

This table shows the average male have higher death rates then the females. males being 64 and females 36. 

## Inferential results 

```{r}
chisq.test(table1)
chisqtestGC(table1)
fisher.test(table1)
```

The p-value for the CHI-squared test is 0.0012, since its less than 0.05 I will reject the null hypothesis and say that there is a difference in death rates based on sex. With that the results are significant between the variables with an estimated odds ratio of 2.3 adn 95% confidence interval ranging from 1.4 to 3.9. 

## Conclusions

